Paper figures

### 2019-02-14 gg_topi.Rmd
added panel
Added figures

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- loadQiimeData()

gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))

@@

# Correlation of tumor size at T7 and abundance of Bacteroidales 24/ through timepoints
```{r}
mass_247 <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  mutate(mouseID=as.integer(stringr::str_sub(.$SampleID,8, 10))) %>%
  left_join(mappingfile[mappingfile$timepoint=='T7',], by=c('mouseID' = 'id-topo'))

@@

# Correlation of tumor size at T7 and difference of abundance of Bacteroidales 24/7 through timepoints
```{r}
mass_247 %>%
  filter(timepoint.x %in% c('T3','T7')) %>%
  select(Abundance, timepoint.x, condition, tumor_mass,mouseID) %>%
  tidyr::spread(timepoint.x, Abundance) %>%
  mutate(T3_7 = abs(T7-T3)) %>%
  ggplot(aes(T3_7,tumor_mass)) +
  geom_point() +
  geom_smooth(method = 'lm')
```

>>>>

```{r load qiime, include=FALSE, cache=TRUE}
gg_topi <- loadQiimeData()
mappingfile <- data.frame(sample_data(gg_topi))
gg_topi.genus <- tax_glom(gg_topi,taxrank = "Genus")
gg_topi.genus.relab <- transform_sample_counts(gg_topi.genus, function(x) x/sum(x))

@@

# Correlation of tumor size at T7 and abundance of Bacteroidales 24/ through timepoints
```{r}
mass_247 <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  mutate(mouseID=as.integer(stringr::str_sub(.$SampleID,8, 10))) %>%
  mutate(mouseID=as.factor(mouseID))%>%
  left_join(mappingfile[mappingfile$timepoint=='T7',], by=c('mouseID' = 'id.topo'))

@@

# Correlation of tumor size at T7 and difference of abundance of Bacteroidales 24/7 through timepoints
```{r}
mass_247 %>%
  filter(timepoint.x %in% c('T3','T5','T6','T7')) %>% 
  select(Abundance, timepoint.x, condition, tumor_mass,mouseID) %>%
  tidyr::spread(timepoint.x, Abundance) %>%
  mutate(T3_7 = abs(T7-T3), T3_6 = abs(T6-T3),T3_5 = abs(T5-T3)) %>%
  select(-T3,-T5,-T6,-T7) %>%
  tidyr::gather(key = timepoint, value=abundance, -condition, -tumor_mass, -mouseID) %>%
  ggplot(aes(abundance, tumor_mass)) +
  geom_point() +
  facet_grid(timepoint~condition)+
  geom_smooth(method = 'lm') + 
  theme_cm() +

### 2019-02-14 paper_figures.Rmd
Added figures

```{r fig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3', fig.height=3, fig.width=3}
annot_df <- (ttest_res %>% filter(Measure == 'Chao1') %>% filter(pvalue < 0.05) %>% mutate(pvalue = paste('p=',signif(pvalue,2), sep = '')))
filt_alpha %>%
  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Chao1')) %>%
  ggplot(aes(timepoint, value)) +
  geom_boxplot(aes(fill=timepoint), width=0.8, alpha=.8) +
  xlab('Timepoint') +
  ylab('Chao1 Index') +
  scale_fill_manual(name = "Timepoint", values = cm_palette) +
  # ggsignif::geom_signif(comparisons = list(c('T1','T3')), y_position = 240, tip_length = 0, test = 't.test') +
  # ggsignif::geom_signif(comparisons = list(c('T2','T3')), y_position = 230, tip_length = 0, test = 't.test') +
  geom_signif(data = annot_df, manual=TRUE, aes(annotations = pvalue, xmin=Timepoint1, xmax=Timepoint2, y_position = c(235,225))) +
  guides(fill=FALSE) +
  theme_cm()
```

@@

```{r fig3b, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3", fig.dim = c(6,5)}
subset(ordwu, timepoint %in% c('T2','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_manual(name = "Timepoint", values = cm_palette[c(2,3)]) +
  theme_cm() +
  annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))),
    gp=grid::gpar(col="black", fontsize=11)),  xmin = 0.15, xmax = .15, ymin = .12, ymax = .12)
```

@@

```{r fig3c, fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.", fig.dim=c(6,5)}
annot_data <- res.filt_genus[unique(lacno_bact_relabs$OTU),] %>% 
              data.frame %>% 
              rownames_to_column('OTU') %>% 
              bind_cols(taxlab = c('Lachnospiraceae NK4A136','Bacteroidales S24-7'),  xmin = c('T2','T2'), timepoint = c('T3','T3'), y_position =  c(0.6,0.7)) %>%
              select(annotations = padj, taxlab, timepoint, xmin, y_position)

ggplot(data=lacno_bact_relabs, aes(x=timepoint, y=Abundance)) +
  stat_boxplot(aes(fill=timepoint), alpha=.8) +
  facet_wrap(~taxlab) +
  scale_fill_manual('Timepoint', values=cm_palette[c(2,3)]) + 
  geom_signif(data = annot_data,
              aes(annotations = signif(annotations,2), xmin=xmin, xmax=timepoint, y_position = y_position), manual = TRUE) +
  theme_cm() +
  theme(strip.text = element_text(size=11)) +
  xlab('Timepoint') +
  ylab('Relative abundance')+
  guides(fill=FALSE)
```

```{r fig3d, fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints", fig.dim = c(6,5)}
not12 <- gg_topi %>%
  subset_samples(!(timepoint %in% c("T1", "T2")))

@@

centroids <- aggregate(cbind(Axis.1, Axis.2)~timepoint, x, mean)
centroids[1:4,c('X2','Y2')] <- centroids[-1,c(2:3)]
ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
  geom_point(size = 1.5, alpha=0.7) +
  geom_point(data=centroids, size = 4) + 
  geom_segment(aes(x=Axis.1, y=Axis.2, xend=X2, yend=Y2), 
               arrow = arrow(length = unit(0.1,'inches')),
               data=centroids, group="1") + 
  scale_color_manual(name = "Timepoint", values = cm_palette[-c(1,2)]) +
  theme_cm()
```

## Panel ?: Ordination beta diversity T3 vs T7
```{r fig3x, fig.dim = c(6, 5)}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T7','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_manual(name = "Timepoint", values = cm_palette[c(3,7)])+
  theme_minimal() +
  annotation_custom(grob = grid::textGrob(bquote(atop(italic("p-value")==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=11)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
```

@@

```{r fig3e, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered", fig.height=6}
genus_ab_t %>%
  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  filter(OTU %in% (genus_ab_t %>%
  group_by(OTU, timepoint) %>%
  summarise(median=median(Abundance)) %>%
  group_by(OTU) %>%
  summarise(filt=min(median)>0.01) %>% filter(filt==TRUE))$OTU) %>%
  filter(Phylum != "Unassigned") %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown'), paste('Unknown',Family,sep = ' '), paste(taxlab))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown Unknown'), paste('Unknown',Phylum,sep = ' '), paste(taxlab))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Uncultured uncultured'), paste('Uncultured',Phylum,sep = ' '), paste(taxlab))) %>%
ggplot() +
  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85),alpha=.8) +
  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
  scale_fill_manual(name = "Timepoint", values = cm_palette[c(3,7)]) +
  coord_flip() +
  xlab('') +
  ylab("Relative abundance") +
  theme_cm()
```

```{r supplfig3e, fig.cap="Fig 3E: Boxplot of abundance of significant genera from DESeq2 differential analysis", fig.height=9}
genus_ab_t %>%
  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  filter(Phylum != "Unassigned") %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown'), paste('Unknown',Family,sep = ' '), paste(taxlab))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown Unknown'), paste('Unknown',Phylum,sep = ' '), paste(taxlab))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Uncultured uncultured'), paste('Uncultured',Phylum,sep = ' '), paste(taxlab))) %>%
ggplot() +
  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85),alpha=.8) +
  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
  scale_fill_manual(name = "Timepoint", values = cm_palette[c(3,7)]) +
  coord_flip() +
  scale_y_log10() +
  xlab('') +
  ylab("Relative abundance") +
  theme_cm() +
  theme(strip.text.y = element_text(angle = 360),
        legend.position = "bottom")
```

@@

```{r fig3f, fig.cap="Fig 3 F: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment"}
idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")
idx_lachno <- which(tax_table(gg_topi.genus.relab)[,'Genus'] == "Lachnospiraceae NK4A136 group" )

@@

otu_table(gg_topi.genus.relab)[c(idx_247,idx_lachno),] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
  mutate(taxlab = str_replace(ifelse(stringr::str_detect(Genus,'uncultured'), paste(Family,sep = ''), paste(Genus, sep = '')), ' group','')) %>%
ggplot() +
  geom_boxplot(aes(timepoint, Abundance, fill=condition), alpha=.8) +
  scale_fill_manual(name = "Treatment", values = c("#a6d854","#ffd92f","#66c2a5","#fc8d62")) +
  facet_wrap(facets = 'taxlab', nrow = 2, scales = "free_x") +
  geom_signif(data = annot_df, manual = TRUE, aes(xmin = xmin, xmax = xmax, annotations = annotations, y_position = y_position),textsize=3) +
  theme_cm() +
  ylab('Relative abundance') +
  xlab('Timepoint')
```

@@

```{r fig3g, include=FALSE, eval=FALSE, fig.height=3,fig.cap="Fig 3 G: Scatterplot of tumor size and difference of beta diversity "}
distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  mutate(timepoint = paste('T',timepoint2,'-T3',sep = '')) %>%
  ggplot(aes(value,tumor_mass)) +
  geom_point() +
  facet_wrap(facets = 'timepoint', nrow = 1, shrink = TRUE, scales = 'free_y') +
  geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
  ylab(bquote('Tumor mass'~( mm^3))) +
  xlab('Bray Curtis distance') +
  theme_cm() +
  geom_text(data = ds_corr_braycf, aes(x=x,y=y, label=label))
```

```{r fig3i, fig.cap="Fig 3I: Scatterplot of tumor size and difference of beta diversity ", fig.height=3}
tbl_data <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
  data.frame %>%
  tidyr::gather_("SampleID", "Abundance", colnames(.)) %>%
  mutate(mouseid=stringr::str_sub(.$SampleID,8, 10)) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  filter(timepoint %in% c('T3','T7')) %>%
  select(-SampleID) %>%
  spread(timepoint,Abundance) %>%
  mutate(diff_abb = T3-T7) %>%
  arrange(mouseid) %>%
  bind_cols( gg_topi.genus.relab %>%
              sample_data %>%
              filter(timepoint == 'T7') %>%
              select(id.topo, tumor_mass) %>%
              arrange(id.topo) %>%
              select(-id.topo))

tbl_data %>%
ggplot(aes(diff_abb, tumor_mass)) +
  geom_point() +
  facet_wrap(facets = 'condition', nrow = 1, scales = 'free') +
  geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
  theme_cm() +
  xlab('Relative abundance (T3-T7)') +
  ylab(bquote('Tumor mass'~( mm^3)))
```

>>>>

```{r fig3a, fig.cap='Fig 3A: Alpha diversity controls T1 vs T2 and pre and post challenge T2 vs T3', fig.height=3, fig.width=2}
annot_df <- (ttest_res %>% filter(Measure == 'Chao1') %>% filter(pvalue < 0.05) %>% mutate(pvalue = paste('p=',signif(pvalue,2), sep = '')))
filt_alpha %>%
  melt(id.vars = c('Description','condizione','timepoint'), measure.vars =  c('Chao1')) %>%
  ggplot(aes(timepoint, value)) +
  geom_boxplot(aes(fill=timepoint), width=0.5, alpha=.8) +
  xlab('Timepoint') +
  ylab('Chao1 Index') +
  scale_fill_manual(name = "Timepoint", values = cm_palette) +
  # ggsignif::geom_signif(comparisons = list(c('T1','T3')), y_position = 240, tip_length = 0, test = 't.test') +
  # ggsignif::geom_signif(comparisons = list(c('T2','T3')), y_position = 230, tip_length = 0, test = 't.test') +
  geom_signif(data = annot_df, manual=TRUE, aes(annotations = pvalue, xmin=Timepoint1, xmax=Timepoint2, y_position = c(235,225))) +
  guides(fill=FALSE) +
  theme_cm()
```

@@

```{r fig3b, fig.cap="Fig 3B: Beta diversity differences pre and post challange T2 vs T3", fig.dim = c(5,3)}
subset(ordwu, timepoint %in% c('T2','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_manual(name = "Timepoint", values = cm_palette[c(2,3)]) +
  theme_cm() +
  annotation_custom(grob = grid::textGrob(label = bquote(atop("p-value"==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))),
    gp=grid::gpar(col="black", fontsize=11)),  xmin = 0.15, xmax = .15, ymin = .12, ymax = .12)
```

@@

```{r fig3c, fig.cap="Fig 3C: Differential abundance analysis highlighted Bacteroidales S24/7 and Lacnospiraceae as most differential abunant families.", fig.dim=c(3,3)}
annot_data <- res.filt_genus[unique(lacno_bact_relabs$OTU),] %>% 
              data.frame %>% 
              rownames_to_column('OTU') %>% 
              bind_cols(taxlab = c('Lachnospiraceae NK4A136','Bacteroidales S24-7'),  xmin = c('T2','T2'), timepoint = c('T3','T3'), y_position =  c(0.6,0.7)) %>%
              select(annotations = padj, taxlab, timepoint, xmin, y_position)

ggplot(data=lacno_bact_relabs, aes(x=timepoint, y=Abundance)) +
  stat_boxplot(aes(fill=timepoint), alpha=.8, width=0.5) +
  facet_wrap(~taxlab) +
  scale_fill_manual('Timepoint', values=cm_palette[c(2,3)]) + 
  geom_signif(data = annot_data,
              aes(annotations = signif(annotations,2), xmin=xmin, xmax=timepoint, y_position = y_position), manual = TRUE) +
  theme_cm() +
  theme(strip.text = element_text(size=11)) +
  xlab('Timepoint') +
  ylab('Relative abundance')+
  guides(fill=FALSE)
```

```{r, cache=TRUE}
not12 <- gg_topi %>%
  subset_samples(!(timepoint %in% c("T1", "T2")))

@@

centroids <- aggregate(cbind(Axis.1, Axis.2)~timepoint, x, mean)
centroids[1:4,c('X2','Y2')] <- centroids[-1,c(2:3)]
```

```{r fig3d, fig.cap="Fig 3D: Ordination centroids beta diversity all timepoints", fig.dim = c(5,3)}
ggplot(x, aes(Axis.1, Axis.2, color=timepoint)) +
  geom_point(size = 1.5, alpha=0.7) +
  geom_point(data=centroids, size = 4) + 
  geom_segment(aes(x=Axis.1, y=Axis.2, xend=X2, yend=Y2), 
               arrow = arrow(length = unit(0.1,'inches')),
               data=centroids, group="1") + 
  scale_color_manual(name = "Timepoint", values = cm_palette[-c(1,2)]) +
  theme_cm()
```

## Panel ?: Ordination beta diversity T3 vs T7
```{r fig3x, fig.dim = c(5, 3)}
metadata_2cat <- mappingfile[mappingfile$timepoint %in% c('T7','T3'),]
distwu_2cat <- phyloseq::distance(subset_samples(gg_topi, timepoint %in% c('T7','T3')), "wunifrac")
permanova <- vegan::adonis(formula = distwu_2cat ~ merge_timepoint, data = metadata_2cat)

subset(ordwu, timepoint %in% c('T7','T3')) %>%
ggplot(aes(Axis.1, Axis.2)) +
  geom_point(aes(color=timepoint), size=3, alpha=.8) +
  scale_color_manual(name = "Timepoint", values = cm_palette[c(3,7)])+
  theme_cm() +
  annotation_custom(grob = grid::textGrob(bquote(atop(italic("p-value")==.(permanova$aov.tab$`Pr(>F)`[1]),italic(R^2)==.(signif(permanova$aov.tab$R2[1], 2)))), gp=grid::gpar(col="black", fontsize=11)),  xmin = 0, xmax = 0, ymin = .1, ymax = .1)
```

@@

```{r fig4a, fig.cap="Fig 4A: Boxplot of abundance of significant genera from DESeq2 differential analysis. Genera with median below 1% are filtered", fig.dim = c(5,3)}
genus_ab_t %>%
  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  filter(OTU %in% (genus_ab_t %>%
  group_by(OTU, timepoint) %>%
  summarise(median=median(Abundance)) %>%
  group_by(OTU) %>%
  summarise(filt=min(median)>0.01) %>% filter(filt==TRUE))$OTU) %>%
  filter(Phylum != "Unassigned") %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('',Family,sep = ''), paste(Genus))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown'), paste('Unknown',Family,sep = ' '), paste(taxlab))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown Unknown'), paste('Unknown',Phylum,sep = ' '), paste(taxlab))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Uncultured uncultured'), paste('Uncultured',Phylum,sep = ' '), paste(taxlab))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'group'), str_replace(taxlab, ' group',''), paste(taxlab))) %>%
ggplot() +
  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85),alpha=.8) +
  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
  scale_fill_manual(name = "Timepoint", values = cm_palette[c(3,7)]) +
  coord_flip() +
  xlab('') +
  ylab("Relative abundance") +
  theme_cm()
```

```{r supplfig4a, fig.cap="Supplemenrary Fig 4a: Boxplot of abundance of significant genera from DESeq2 differential analysis", fig.height=9}
genus_ab_t %>%
  bind_cols(data.frame(tax_table(ggtopi_subs.genus.relab)[genus_ab_t$OTU,c('Phylum','Family','Genus')])) %>%
  filter(Phylum != "Unassigned") %>%
  mutate(taxlab = ifelse(stringr::str_detect(Genus,'uncultured'), paste('Uncultured',Family,sep = ' '), paste(Genus))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown'), paste('Unknown',Family,sep = ' '), paste(taxlab))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Unknown Unknown'), paste('Unknown',Phylum,sep = ' '), paste(taxlab))) %>%
  mutate(taxlab = ifelse(stringr::str_detect(taxlab,'Uncultured uncultured'), paste('Uncultured',Phylum,sep = ' '), paste(taxlab))) %>%
ggplot() +
  geom_boxplot(aes(reorder(taxlab, Abundance, FUN = median), Abundance, fill=timepoint),  position = position_dodge(-.85),alpha=.8) +
  facet_grid(Phylum ~ ., scales = "free", space = "free_y", as.table= T) +
  scale_fill_manual(name = "Timepoint", values = cm_palette[c(3,7)]) +
  coord_flip() +
  scale_y_log10() +
  xlab('') +
  ylab("Relative abundance") +
  theme_cm() +
  theme(strip.text.y = element_text(angle = 360),
        legend.position = "bottom")
```

@@

```{r fig4B, fig.cap="Fig 4B: Abundance of Bacteroidales 24/7 and Lachnospiraceae NK4A136 through timepoints and treatment", fig.dim=c(5,4)}
idx_247 <- which(tax_table(gg_topi.genus.relab)[,'Family'] == "Bacteroidales S24-7 group" & tax_table(gg_topi.genus.relab)[,'Genus'] == "uncultured bacterium")
idx_lachno <- which(tax_table(gg_topi.genus.relab)[,'Genus'] == "Lachnospiraceae NK4A136 group" )

@@

otu_table(gg_topi.genus.relab)[c(idx_247,idx_lachno),] %>%
  data.frame %>%
  tibble::rownames_to_column("OTU") %>%
  tidyr::gather_("SampleID", "Abundance", setdiff(colnames(.), "OTU")) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  bind_cols(data.frame(tax_table(gg_topi.genus.relab)[.$OTU,c('Phylum','Family','Genus')])) %>%
  mutate(taxlab = str_replace(ifelse(stringr::str_detect(Genus,'uncultured'), paste(Family,sep = ''), paste(Genus, sep = '')), ' group','')) %>%
ggplot() +
  geom_boxplot(aes(timepoint, Abundance, fill=condition), alpha=.8) +
  scale_fill_manual(name = "Treatment", values = c("#a6d854","#ffd92f","#66c2a5","#fc8d62")) +
  facet_wrap(facets = 'taxlab', nrow = 2, scales = "free_x") +
  geom_signif(data = annot_df, manual = TRUE, aes(xmin = xmin, xmax = xmax, annotations = annotations, y_position = y_position),textsize=3) +
  theme_cm() +
  theme(strip.text = element_text(family = "sans-serif", size = 9, colour = "black"),
        axis.text = element_text(family = "sans-serif", size = 8, colour = "black"),
        axis.title = element_text(family = "sans-serif", size = 8, colour = "black"))+
  ylab('Relative abundance') +
  xlab('Timepoint')
```

@@

```{r fig4C, fig.cap="Fig 4C: Scatterplot of tumor size and difference of beta diversity ", fig.dim=c(5,2)}
distancesbc %>% 
  dplyr::filter(topo1==topo2) %>%
  dplyr::filter(timepoint1==3) %>%
  dplyr::filter(timepoint2>4) %>%
  dplyr::select(cols, timepoint2, value) %>%
  dplyr::left_join(mappingfile, by=c('cols' = 'Description')) %>%
  mutate(timepoint = paste('T',timepoint2,'-T3',sep = '')) %>%
ggplot(aes(value,tumor_mass)) +
  geom_point(size=2) +
  facet_wrap(facets = 'timepoint', nrow = 1, shrink = TRUE, scales = 'free_y') +
  geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
  ylab('Tumor mass (mm3)') +
  xlab('Bray Curtis distance') +
  geom_text(data = ds_corr_braycf, aes(x=x,y=y, label=label), size=2) +
  theme_cm() +
  theme(strip.text = element_text(family = "sans-serif", size = 9, colour = "black"),
        axis.text = element_text(family = "sans-serif", size = 8, colour = "black"),
        axis.title = element_text(family = "sans-serif", size = 8, colour = "black"))
```

```{r}
tbl_data <- otu_table(gg_topi.genus.relab)[idx_247,] %>%
  data.frame %>%
  tidyr::gather_("SampleID", "Abundance", colnames(.)) %>%
  mutate(mouseid=stringr::str_sub(.$SampleID,8, 10)) %>%
  mutate(timepoint=stringr::str_sub(.$SampleID,12, 13)) %>%
  mutate(condition= stringr::str_sub(.$SampleID,4, 6)) %>%
  filter(timepoint %in% c('T3','T7')) %>%
  select(-SampleID) %>%
  spread(timepoint,Abundance) %>%
  mutate(diff_abb = T3-T7) %>%
  arrange(mouseid) %>%
  bind_cols( gg_topi.genus.relab %>%
              sample_data %>%
              filter(timepoint == 'T7') %>%
              select(id.topo, tumor_mass) %>%
              arrange(id.topo) %>%
              select(-id.topo))
```
```{r fig4D, fig.cap="Fig 4D: Scatterplot of tumor size and difference of beta diversity ", fig.dim=c(3.5,3.5)}
tbl_data %>%
ggplot(aes(diff_abb, tumor_mass)) +
  geom_point(size=2) +
  facet_wrap(facets = 'condition', nrow = 2, scales = 'free') +
  geom_line(stat='smooth', method = 'lm', color='red', alpha=0.7, size=.7) + 
  theme_cm() +
  theme(strip.text = element_text(family = "sans-serif", size = 9, colour = "black"),
        axis.text = element_text(family = "sans-serif", size = 8, colour = "black"),
        axis.title = element_text(family = "sans-serif", size = 8, colour = "black")) +
  xlab('Relative abundance (T3-T7)') +
  ylab('Tumor mass (mm3)')
```
